### 2018-03-14 paper/3-abstractions.tex
Finish S3.5

We can now say what it means for a build system to be \emph{correct}, something
that is seldom stated formally:

\definition[Correctness]{Suppose \hs{build} is a build system, \hs{task} is a
build task description, \hs{key} is an output key, \hs{store} is an initial
store, and \hs{result} is the store produced by running the build system with
parameters \hs{task}, \hs{key} and \hs{store}. Or, using the precise language of
our abstractions:

@@

\begin{itemize}
    \item The \hs{ideal} store is \emph{consistent} with the \hs{task}, i.e.
    for all possible non-input keys \hs{k}, the result of executing the
    \hs{task} matches the value in the \hs{ideal} store:
    \[
    \hs{execute}~\hs{task}~\hs{ideal}~\hs{k}~\hs{==}~\hs{Just}~\hs{(}\hs{ideal}~\hs{k)}.
    \]
    \item The \hs{result} and \hs{ideal} agree on the output \hs{key}, i.e.
    \hs{result}~\hs{key}~\hs{==}~\hs{ideal}~\hs{key}. In other words, we require
    that the output \hs{key} has the correct value, but do not impose any
    restrictions on intermediate keys, therefore permitting shallow cloud builds.
    \item The \hs{store}, \hs{ideal} and \hs{result} all agree on the inputs
    that belong to the transitive closure of \hs{key}'s dependencies. This means
    that (i) no inputs were corrupted during the build, and (ii)~the \hs{ideal}
    store has constraints not only on the output, but also on the inputs.
\end{itemize}
}
\label{def-correct}

@@

We can also \hs{execute} a compute on a given store. This relies on a similar
transformation, now based on the \hs{Identity} monad (see
Appendix~\S\ref{sec-appendix-transformers} for the implementation of
\hs{Identity}):

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Data.Functor.Identity
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
execute :: Task Monad k v -> (k -> v) -> k -> Maybe v
execute compute store = fmap runIdentity . compute (pure . store)
\end{minted}
\vspace{1mm}

\noindent
Note that \hs{execute} can accept both an applicative as well as a \emph{monadic
task} introduced in~\S\ref{sec-task-monad}. Below we run \hs{execute} on
the above \hs{add} example. Note that \hs{execute} returns \hs{Nothing} for
inputs.

>>>>

We can now say what it means for a build system to be \emph{correct}, something
that is seldom stated formally. Before proceeding, we need to introduce a
useful utility function called \hs{execute} that executes a \hs{task} on a
\hs{key} by reading its dependencies from a supplied \hs{store}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
newtype Identity a = Identity { runIdentity :: a }
\end{minted}
\vspace{0mm}
\begin{minted}[xleftmargin=10pt]{haskell}
execute :: Task Monad k v -> Store i k v -> k -> Maybe v
execute task store key = runIdentity <$> task (Identity . getValue store) key
\end{minted}
\vspace{1mm}

\noindent
This function is simpler than it might look: it fixes the `impedance mismatch'
between the pure function \hs{getValue}~\hs{store}, whose type is \store, and
the \hs{fetch} argument of the \hs{task}, whose type must be
\hs{k}~\hs{->}~\hs{f}~\hs{v} for some \hs{f}. To fix the mismatch, we wrap the
result of the pure function in the \hs{Identity} monad: the function
\hs{Identity}~\hs{.}~\hs{getValue}~\hs{store} has the type
\hs{k}~\hs{->}~\hs{Identity}~\hs{v}, and can now be passed to a \hs{task}. The
result comes as \hs{Maybe}~\hs{(Identity}~\hs{v)}, hence we now need to get rid
of the \hs{Identity} wrapper by applying \hs{runIdentity} to the contents of
\hs{Maybe}.

\vspace{-1mm}
\noindent\rule{\textwidth}{0.4pt}
\vspace{-7mm}
\definition[Correctness]{Suppose \hs{build} is a build system, \hs{task} is a
build task description, \hs{key} is an output key, \hs{store} is an initial
store, and \hs{result} is the store produced by running the build system with
parameters \hs{task}, \hs{key} and \hs{store}. Or, using the precise language of
our abstractions:

@@

\begin{itemize}
    \item The \hs{ideal} store is \emph{consistent} with the \hs{task}, i.e.
    for all possible non-input keys \hs{k}, the result of executing the
    \hs{task} matches the value in the \hs{ideal} store:
    \vspace{-1mm}
    \[
    \hs{execute}~\hs{task}~\hs{ideal}~\hs{k}~\hs{==}~\hs{Just}~\hs{(}\hs{getValue}~\hs{ideal}~\hs{k)}.
    \]
    \item The \hs{result} and \hs{ideal} agree on the output \hs{key}, i.e.:
    \vspace{-1mm}
    \[
        \hs{getValue}~\hs{result}~\hs{key}~\hs{==}~\hs{getValue}~\hs{ideal}~\hs{key}.
    \]
    In other words, we require that the output \hs{key} has the correct value,
    but do not impose any restrictions on intermediate keys, therefore
    permitting shallow cloud builds.
    \item The \hs{store}, \hs{result} and \hs{ideal} all agree on the inputs
    that belong to the transitive closure of \hs{key}'s dependencies:
    \[
        \hs{and}~\hs{[}~\hs{getValue}~\hs{store}~\hs{k}~\hs{==}~\hs{getValue}~\hs{result}~\hs{k}~\hs{|}~\hs{k}~\hs{<-}~\hs{inputs}~\hs{key}~\hs{]}
    \]
    \vspace{-6mm}
    \[
        \hs{and}~\hs{[}~\hs{getValue}~\hs{ideal}~\hs{k}~\hs{==}~\hs{getValue}~\hs{result}~\hs{k}~\hs{|}~\hs{k}~\hs{<-}~\hs{inputs}~\hs{key}~\hs{]}
    \]
    where the implementation of \hs{inputs} will be provided in the next
    subsection~\S\ref{sec-deps}, which is dedicated to the task dependency
    analysis. This requirement guarantees that (i) no inputs were corrupted
    during the build, and (ii)~the \hs{ideal} store has constraints not only
    on the output, but also on the inputs.
\end{itemize}
A build system is \emph{correct} if it produces a correct \hs{result} for any
given \hs{task}, \hs{key} and \hs{store}.
}
\label{def-correct}
\vspace{-2mm}
\rule{\textwidth}{0.4pt}

It is hard to satisfy the above definition of correctness given a task
description with cycles. All build systems discussed in this paper are correct
only under the assumption that the given task description is acyclic. This
includes the \hs{busy} build system introduced earlier: it will loop
indefinitely given a cyclic \hs{task}. Some build systems provide a limited
support for cyclic tasks, see~\S\ref{sec-iterative-compute}.

@@
